---
title: "Data Cleaning for Marketing Data"
author: "Maya Reese Farmer"
date: "October 19, 2021"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

#### Introduction  
These were the data cleaning techniques I used on the Marketing Analytics dataset from kaggle.com  
This data was originally provided by Dr. Omar Romero-Hernandez.

The marketing dataset contains 2240 customers and has 28 columns with data on:  

 - customer profiles (ID, birth year, education, income, etc.) 
 - product preferences
 - marketing campaign successes/failures
 - channel performance


```{r}
## importing dataset into r
marketing_data <- read.csv("~/Downloads/marketing_data.csv")

## view data
head(marketing_data,10)
```
 
 
#### Data Cleaning

```{r}
## task 1: convert income to numerical value
library(readr)
library(dplyr)

income = marketing_data %>% select(Income)
income_parsed = sapply(income, parse_number) %>% as_tibble()
clean.marketing = marketing_data %>% select(-Income) %>% bind_cols(income_parsed)
```

```{r}
## task 2: create an age column from year_birth
clean.marketing$age = (2021 - clean.marketing$Year_Birth)
```

```{r}
## task 3: create age group categories
clean.marketing = clean.marketing %>%
  mutate(AgeGroup = case_when(
    clean.marketing$age <= 35 ~ "Under 36",
    clean.marketing$age >= 36 & clean.marketing$age <= 50 ~ "36-50",
    clean.marketing$age >= 51 & clean.marketing$age <= 65 ~ "51-65",
    clean.marketing$age >= 66 & clean.marketing$age <= 80 ~ "66-80",
    clean.marketing$age >= 81 ~ "80+"))
```

```{r}
## task 4: create income range categories
clean.marketing = clean.marketing %>%
  mutate(IncomeRange = case_when(
    clean.marketing$Income < 50000 ~ "Under 50K",
    clean.marketing$Income >= 50000 & clean.marketing$Income <= 75000 ~ "50K-75K",
    clean.marketing$Income > 75000 & clean.marketing$Income <= 100000 ~ "75K-100K",
    clean.marketing$Income > 100000 ~ "Over 100K"))
```

```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
## task 5: rename some columns (this makes things easier to categorize when we pivot the table later)
clean.marketing = clean.marketing %>% rename(Web = NumWebPurchases, Catalog = NumCatalogPurchases, Store = NumStorePurchases,Camp1 = AcceptedCmp1,Camp2 = AcceptedCmp2,Camp3 = AcceptedCmp3,Camp4 = AcceptedCmp4,Camp5 = AcceptedCmp5)
```

```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
## task 6: create column that sums the total amount each customer has spent
clean.marketing = clean.marketing %>% mutate(total.spent = rowSums(clean.marketing[,9:14]))
```

---
